---
title: "Need for Linear mixed effect model"
author: "Chandan Pandey"
date: "19/10/2022"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# When do we need GLMM.

- Some data are very messy and complex. 
- eg., effect of $N_{2}$ on grass growth; Many factor can influence
  - Species difference. 
  - Local Area variable. 
  - Species composition. 
  - Other Nutrient factors. 
  - And many many more. 
- Impossible to account for all the predictors.



# Simpson's Paradox

Let us check from the live example from the simulated dataset. 
```{r simpsons paradox,echo=FALSE}
group1_x <- runif(100,1,10)
group_2_x <- runif(100,5,15)
group3_x <-runif(100,18,30)
group1_y <- rnorm(100,10,2)*group1_x+150+rnorm(100,0,3)
group2_y <- rnorm(100,10,2)*group_2_x+100+rnorm(100,0,3)
group3_y <- rnorm(100,10,2)*group_2_x+50+rnorm(100,0,3)
group<-c(rep("A",100),rep("B",100),rep("C",100))
simps_para<-data.frame(x=c(group1_x,group_2_x,group3_x),
                       y=c(group1_y,group2_y,group3_y),
                       group)
simps_para$group<-factor(simps_para$group)
plot(y~x,data = simps_para)
```

#load("C:/Users/chand/Documents/R-advance/Advance-R/data/dragons.RData")
- Bullet 1
- Bullet 2
- Bullet 3

# Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

