---
title: "Need for Linear mixed effect model"
author: "Chandan Pandey"
date: "19/10/2022"
output:
  beamer_presentation: default
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits = 3)
```

# When do we need GLMM.

- Some data are very messy and complex. 
- eg., effect of $N_{2}$ on grass growth; Many factor can influence
  - Species difference. 
  - Local Area variable. 
  - Species composition. 
  - Other Nutrient factors. 
  - And many many more. 
- Impossible to account for all the predictors.



# Simpson's Paradox

What is your intuition about relation between X and Y ?


```{r simpsons paradox,echo=FALSE,warning=F,include=T}
library(nlme)
group1_x <- runif(100,10,30)
group_2_x <- runif(100,20,40)
group3_x <-runif(100,30,50)
group1_y <- rnorm(100,5,1)*group1_x+rnorm(100,300,10)+rnorm(100,0,3)
group2_y <- rnorm(100,5,1)*group_2_x+rnorm(100,200,10)+rnorm(100,0,3)
group3_y <- rnorm(100,5,1)*group3_x+rnorm(100,100,10)+rnorm(100,0,3)
group<-c(rep("A",100),rep("B",100),rep("C",100))
simps_para<-data.frame(x=c(group1_x,group_2_x,group3_x),
                       y=c(group1_y,group2_y,group3_y),
                       group)
simps_para$group<-factor(simps_para$group)
plot(y~x,data = simps_para)
mod_simple_lm <- lm(y~x,data = simps_para)
library(dplyr)

```



# Let plot this model with simple linear model



```{r simple lm, echo=F,warning=FALSE,fig.align='left'}

mod_simple_lm <- lm(y~x,data = simps_para)
summary1<-summary(mod_simple_lm)[4]%>%as.data.frame()
names(summary1) <- c("Estimate","SE","t_value","P")
print(summary1)
```

# Let plot this model with simple linear model

```{r simple model val, echo=F}
##model validation
par(mfrow=c(2,2))
plot(mod_simple_lm)
```

# Let plot this model with simple linear model

```{r simple model plot,echo=FALSE}
plot(y~x,data = simps_para)
abline(mod_simple_lm,lwd=4)
```

# Now let me group these variable 

## I have lied about dataset here a bit

These value belongs to 3 groups. Think these groups as

- Three different species
- Three trials
- Plots
- communities
- subspecies 
- and many more

__What will say on relation of x and Y ?__

# Here is plot

```{r plot with group, echo = F,fig.align='right',fig.dim=c(5,5)}
plot(y~x,data = simps_para,col=group)
```

# One way to address this problem

Model with two predictor but there is problem with that as well. We will discuss this latter but first model with two predictor. 

```{r two predictor, echo=F}
mod_simple_2 <- lm(y~x+group,data = simps_para)
summary.aov(mod_simple_2)%>%print()
mod_simple_2$coefficients
```

# Model validation

```{r simple 2 pred mod val, echo=F}
par(mfrow=c(2,2))
plot(mod_simple_2)
```

# Models performance

Let us compare the two models. 

```{r compare the model,echo=FALSE}
## AIC value 
library(AICcmodavg)
aictab(cand.set = list(mod_simple_lm,mod_simple_2),modnames = c("Single","grouped"))
# R squared value
R_1 <- summary(mod_simple_lm)[9]
R_2 <- summary(mod_simple_2)[9]
data.frame(model = c("Single","grouped"),R_squared= c(R_1[[1]],R_2[[1]]))%>%print()
```

Model with two predictor have improved the model but what is the trend

# Plot the trend with two predictors

```{r plot with two trend, echo=F,fig.dim=c(3,1.5)}
library(ggplot2)
ggiraphExtra::ggPredict(mod_simple_2,se=TRUE,interactive=F)
```

Q: 

* What is the correct trend ?
* What was real question this case ? can we generalized this model.

# What is the answer

## Linear Mixed effect model 

# Let model with LMM (1)

```{r LMM, echo=T,warning=F,fig.dim=c(3,2)}
mod_LM <- lme(y~x, random =~1|group,
              data = simps_para,method = "REML")
plot(mod_LM)
```

# Let model with LMM (2)
```{r qq plot lme,echo=FALSE}
qqnorm(mod_LM)
```

# Model summary

```{r summary,echo=FALSE}
summary(mod_LM)
```

# Ploting the model 

```{r ploting LM,echo=TRUE,fig.dim=c(4,2.5)}
plot(y~x,data=simps_para)
abline(a=mod_LM$coefficients$fixed[1],
       b=mod_LM$coefficients$fixed[2],
       col="red",lwd=5)
```