---
title: "Linear Mixed effect model-day2"
author: "Chandan Kumar Pandey"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE,echo=FALSE}
require(ggplot2)
require(GGally)
require(reshape2)
require(lme4)
require(compiler)
require(parallel)
require(boot)
require(lattice)
library(dplyr)
```

# Mixed effect models



# possun Morphometic dataset. 

```{r load the data,echo=TRUE}
possum <- read.csv("./data/possum.csv",header = T)
## Basic data data exploration 
summary(possum)
## convert the population,site and gender  into factors
possum$site <-as.factor(possum$site)
possum$sex <- as.factor(possum$sex)
possum$Pop <-as.factor(possum$Pop)
summary(possum)
## you can see the difference in the way population,site and gender columns
## box plot of lenght vs gender and pop 
boxplot(totlngth~sex+Pop, data = possum)
## pair plot to see if there is co - linearity
ggpairs(possum[,c(8,7,9,10)])
# 
```

# Droping the co-linear variable. 

There is small amount of correlation between tail and skull, foot len vs skull. However, since we want to know of skull width can predict the possum length. 

```{r droping the predictor, echo=T}
#scatter plot
scat_plot <- ggplot(data = possum, aes(x=skullw,y=totlngth,col=sex))
scat_plot + geom_point()+facet_wrap(.~Pop+site)
```

* What you take out of this graphs on the population trend of the possum

# Let fit simple linear model

$$Lenght = \alpha + \beta_{1}* Skull width+ \epsilon $$
Where $$\epsilon \sim N(0,\sigma^{2})$$

```{r pressure, echo=T}
mod_lm <- lm(totlngth~skullw,data = possum)
summary(mod_lm)
par(mfrow=c(2,2))
plot(mod_lm)
```

The data seems to be heterogeneous due to fact from the first graph

Let take population as random effect 

```{r pop as random effect, echo=TRUE}
model_LMM <- lmer(totlngth~skullw + (1|Pop:site),data = possum)
plot(model_LMM)
## clearly intercet are different. 
dotplot(ranef(model_LMM, which = "Pop", condVar = TRUE))
## let us see the model
print(model_LMM, corr = FALSE)
## ##getting SE and tabular format for fixed effect.
se <- sqrt(diag(vcov(model_LMM)))
# table of estimates with 95% CI
(tab <- cbind(Est = fixef(model_LMM), 
              LL = fixef(model_LMM) - 1.96 * se,
              UL = fixef(model_LMM) + 1.96 *se))
```