---
title: "Linear Mixed effect model-day2"
author: "Chandan Kumar Pandey"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE,echo=FALSE}
require(ggplot2)
library(ggiraph)
require(ggiraphExtra)
require(GGally)
require(ggExtra)
require(reshape2)
require(lme4)
library(dplyr)
library(ggeffects)
```

# Mixed effect models



# possun Morphometic dataset. 

```{r load the data,echo=TRUE}
possum <- read.csv("./data/possum.csv",header = T)
## Basic data data exploration 
summary(possum)
## convert the population,site and gender  into factors
possum$site <-as.factor(possum$site)
possum$sex <- as.factor(possum$sex)
possum$Pop <-as.factor(possum$Pop)
summary(possum)
## you can see the difference in the way population,site and gender columns
## box plot of lenght vs gender and pop 
boxplot(totlngth~Pop, data = possum)
boxplot(totlngth~site, data = possum)
## pair plot to see if there is co - linearity
ggpairs(possum[,c(8,7,9,10)])
# 
```

# Droping the co-linear variable. 

There is small amount of correlation between tail and skull, foot len vs skull. However, since we want to know of skull width can predict the possum length. 


* What you take out of this graphs on the population trend of the possum

# Let fit simple linear model

$$Lenght = \alpha + \beta_{1}* Skull width+ \epsilon $$
Where $$\epsilon \sim N(0,\sigma^{2})$$

```{r pressure, echo=T}
mod_lm <- glm(totlngth~skullw,data = possum,family = "gaussian")
summary(mod_lm)
par(mfrow=c(2,2))
plot(mod_lm)
ggPredict(mod_lm,se = T,interactive = F)+labs(x="Skull Width", y= "Body Length")
```
Clearly there the data is not homogeneous,

* We need to transform the variable to see if we can get out model to work

## New model will be 

$$\log(Length) = \alpha + \beta_{1}* \log(Skull width)+ \epsilon $$

```{r log lenght, echo=T}
possum <- possum%>%mutate(log_skullwid=log(skullw))
possum <- possum%>%mutate(log_len=log(totlngth))
mod_lm_log <- glm(log_len~log_skullwid,data = possum,family = "gaussian")
par(mfrow=c(2,2))
plot(mod_lm_log)
ggPredict(mod_lm_log,se = T,interactive = F)+labs(x="Skull Width (log)", y= "Body Length (log)")
```

What is the assumption we are violating here for linear model of Gaussian Family ?

* Homoscedasticity: The variance of residual is the same for any value of X.
* what about "Independence: Observations are independent of each other."

Let us check

```{r independence check,echo=TRUE}
ggplot(possum, aes(x = site, y = skullw)) +
  geom_jitter(alpha = .1) +
  geom_violin(alpha = 0.7,aes(fill=Pop),draw_quantiles = c(0.25, 0.5, 0.75)) +
  ggtitle("Violin plot of Skull Width vs Site")
  
#scatter plot
scat_plot <- ggplot(data = possum, aes(x=skullw,y=totlngth))
scat_plot + geom_point()+facet_wrap(.~Pop+site)  
# sample size
group_by(possum,site)%>%summarise(count=n())
```

* __the data is not independent__


As we can clearly see from the data collected the data is not independent but nested in the nature. 

Moreover, the data is taken from 2 different population. In addition there are several site nested for each population. 

This is called nested data. 

## One traditional way (wrong way)

Let make model of all the site individually. 

```{r model with each site,echo=TRUE}
site<-unique(possum$site)
Beta <- vector(length = length(site))
alpha <- vector(length = length(site))
for(k in 1:length(Beta)){
  flag_mod <- lm(totlngth~skullw,data = possum[possum$site==site[k],])
  alpha[k] <- as.numeric(flag_mod$coefficients[[1]])
  Beta[k] <- as.numeric(flag_mod$coefficients[[2]])
}
##let plot regression for each site. 
ggplot(possum,aes(x=skullw,y=totlngth))+geom_blank()+
  geom_abline(slope = Beta[1],intercept = alpha[1],col="red")+
  geom_abline(slope = Beta[2],intercept = alpha[2],col="black")+
  geom_abline(slope = Beta[3],intercept = alpha[3],col="darkgreen")+
  geom_abline(slope = Beta[4],intercept = alpha[4],col="khaki4")+
  geom_abline(slope = Beta[5],intercept = alpha[5],col="orange")+
  geom_abline(slope = Beta[6],intercept = alpha[6],col="blue")+
  geom_abline(slope = Beta[7],intercept = alpha[7],col="purple")
```
# Less traditional way (method in this case)

It to account for variation in site. 

```{r account for variation, echo=T}
mod_account_site <- lm(totlngth~skullw + site ,data = possum)
AIC(mod_account_site,mod_lm)
#AIC value have decrease but What is interpretation of this model ?
pred_value <- as.data.frame(predict(mod_account_site,possum,se.fit = T))
pred_value <- cbind(pred_value,possum)
ggplot(data = pred_value,aes(x=skullw,y=totlngth))+
  geom_point(aes(col=site))+
  geom_line(aes(x=skullw,y = fit,group=site))
```
Now it is clear that for some random reason the behavior of data is dependent on site.

# How to solve such problem

* One problem in using model $y = \alpha + \beta*X1 + \beta*X2$ is 

```{r pop as random effect, echo=TRUE}

model_LMM <- lmer(totlngth ~ skullw + (1|site) ,data = possum[is.na(possum$age)==F,])
summary(model_LMM)
## ##getting SE and tabular format for fixed effect.
plot(model_LMM,site~ resid(.), abline = 0 )
plot(model_LMM, resid(., type = "pearson") ~ fitted(.) |site, id = 0.05, 
     adj = -0.3, pch = 20, col = "gray40")
pred <- ggpredict(model_LMM,terms = "skullw")
ggplot(pred) + 
   geom_line(aes(x = x, y = predicted)) +          # slope
   geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error), 
               fill = "lightgrey", alpha = 0.5) +  # error band
   geom_point(data = possum,                      # adding the raw data (scaled values)
              aes(x = skullw, y = totlngth, colour = site)) + 
   labs(x = "Skull width", y = "Total length", 
        title = "Morphometric relation between skull width and body length") + 
   theme_minimal()+geom_abline(slope = 0.72,intercept = 45.63,col="blue")

```

